---
title: "R Markdown Demo"
author: "BUSN 32100"
date: "Oct 5 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

In this demo we continue to work with the starwars data 

##Packages
We start with loading the packages we'll use.
```{r load-packages, message=FALSE ,warning=F}

#1
#a)
library(tidyverse)
state.df = data.frame(state.x77, Region=state.region, Division=state.division)
summary(state.df)
state.df$Abbr<-state.abb
print(state.df[1:3,1:11])
#b)
print(state.df[1:3,-c(9)])
#Region gone!!
#c)
state.df$Center.x<-state.center$x
state.df$Center.y<-state.center$y
print(state.df[1:3,1:12])
#d)
longless <- state.df[state.df$Center.x < -100,]  
longless.data<- data.frame(longless)
#e
longless9 <- data_frame(state.df$Center.x < -100 & state.df$Murder > 9)
which(longless9==TRUE)
#[1]  2  5 28 31
#these four states meet the -100 longitude +9% murder rate criteria
badstates<-which(longless9==TRUE)
state.df$Life.Exp[badstates]
# 69.31 71.71 69.03 70.32
#State #5 has the highest life expectancy. 


```

First, let's try to `knit` this R Markdown file to a html file. Click on the `knit` button on top. It should open a html page.

In class, we have seen these two plots. Now each member of your group can choose one block and modified the code with in that block to show a plot with only two genders, male and female. You can also choose between `knit to html` or `knit to pdf`. Compare your knitted html/pdf with your other member of your group. Let us know if you encounter any error.

===========
```{r}
rio = read.csv("https://raw.githubusercontent.com/BUSN32100/data_files/master/rio.csv")
#2a. What kind of object is rio? What are its dimensions and columns names of rio? What does each row represent? Is there any missing data?
summary(rio)
#Looks like rio includes information on all the athletes participating in the Rio olymipics.
typeof(rio)
dim(rio)
#[1] 11538    12
#each row is an individual athlete. 
is.null(rio)
#FALSE, looks like there is no missing data. 
#2.b) 
#11,538 athlets competed in the olympics
class(rio$nationality) 
levels(rio$nationality) 
summary(rio$nationality)
#The US has the most athletes at 567
summary(rio$sex)
#female   male 
#  5205   6333 
subset.Male <- rio[rio$sex == 'male',] 
#Brazil has the most male athlets at 269 
summary(subset.Male$nationality)
subset.Female <- rio[rio$sex == 'female',] 
#US has the most female athlets at 303. 
summary(subset.Female$nationality)

#Challenge)
summary(rio$name)
sum(summary(rio$name) > 1) 
#22 people share the same name 

#2c) 
sum(rio$gold)
#[1] 666
sum(rio$silver)
# [1] 655
sum(rio$bronze)
# [1] 704
666+655+704
#[1] 2025
#Surpirsingly, there were different amounts of each medal awarded. I did my research and apparently this is because if there is a tie for gold, no silver is awarded. 
rio <- data.frame (rio, total = rio$gold + rio$silver + rio$bronze)
max(rio$total) #The athlete with most medals got 6 (Michael Phelps) 
rio[rio$total == 6,]
```

2. Segmented bar plots, proportions
===========
```{R}
#3a)
install.packages("ggplot2")
library("ggplot2")
ggplot(data=rio, mapping=aes(x=total))+geom_histogram(color = 'red', fill='white')
#b)
rio2 <- data.frame(rio, age = 2016 - rio$year_of_birth)
max(rio2$age)
#[1] 62
rio[rio2$age == 62,] 
#Julie Brougham from New Zealand and Mary Hanna from Australia are the oldest. 
min(rio2$age) #min age is 14
rio[rio2$age == 14,] #there are 8 14 year olds,  7 female and 1 male, all participated in aquatics
ggplot(data=rio2, mapping=aes(x=age, fill=sex))+geom_histogram(bindwidth=10)  
rio3 <- rio2[rio2$gold>0,]
max(rio3$age)
#[1] 59, The max age is 59, Nick Skelton from Great Britain
min(rio3$age)
#[1] 15 Minimum age is 15, Qian Ren from China
ggplot(data=rio3, mapping=aes(x=age, fill=sex))+geom_histogram(bindwidth=10) 
#Looks like the athletes who won a medal tend to be younger 
summary(rio3)
#Gold medalists are on average 26.97 years old 
#I think age is not much of a factor
summary(rio)
#I'm noticing that a disporportionate number of Americans win the gold medal. 
#d)
ggplot(data=rio2, mapping = aes(x=age))+geom_histogram()+facet_wrap(.~sport)
range(rio2[rio2$sport=='golf',]$age)
#[1] 18 47 pretty big age range 
range(rio2[rio2$sport=='gymnastics',]$age)
#[1] 16 41, slightly narrower range than golf
range(rio2[rio2$sport=='rowing',]$age)
#[1] 18 57, more older people
#Makes sense since rowing isn't as physically taxing as gymnastics

#e)
ggplot(data=rio, mapping=aes(y=height, x=sport))+geom_boxplot() +theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggplot(data=rio, mapping=aes(y=weight, x=sport))+geom_boxplot() +theme(axis.text.x = element_text(angle = 90, hjust = 1))

#f) 
ggplot(data=rio, mapping=aes(x=height, y=weight, color=sex))+geom_point(alpha = 0.1) + facet_wrap(.~sport)

#Challenge) 
counts <- summary(rio$nationality)
counts <- as.data.frame (counts)
Country <- row.names(counts)

counts <- data.frame("total_athlets"=counts,Country)

colnames(counts)[1] <- "total_athlets"

ggplot(data=counts[1:50,], mapping=aes(x=total_athlets, y=Country))+geom_point()

```


3. Segmented counts/proportions plots
===========

```

4. Knit to PDF
=========
To be able to knit PDF file from Rmarkdown. Try run the following lines. You will see a warning message saying two error messages will be displayed. Acknowledge it. After two error messages and a long installation process, you should have the ability to generate PDF from Rmarkdown files.
**You may want to do this after the class today, you will need the Rstudio to complete the group lab**

To check if you have install successfully. Run these lines in your console. You should see an output of 
`[1] "test.pdf"` and a `test.pdf` now in your folder.

```{r}
